---
title: "Brownlows Bets"
author: "Erika Amado"
date: "put the date here"
output: html_document: default
 ---

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 1137 times so far. Where do you think his profits are? Discuss this in a markdown document

# Methods

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 1137 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->


#Profit Estimate

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting!-->

First we create the year data and a vector to contain.
```{r}
Bdays = 1:365
Bsamples = vector("numeric")
TotalDollars= 0
saveDollars=vector('numeric')
```

Run this simulation 1137 times 

```{r}
for(loop in 1:1000){
  
  TotalDollars= 0
  classSize= 35
  for(i in 1:1137) {
    Bsamples = sample(Bdays,35, replace= TRUE)
   if (length(unique(Bsamples)) == classSize) {  
     TotalDollars = TotalDollars -2
   } else{
     TotalDollars = TotalDollars +1
   }
  }
  
  saveDollars[loop] = TotalDollars
  }
  hist(saveDollars)
  var(saveDollars)
  print(paste("average return =",mean(saveDollars)))
```

The mean for the whole test is 505.
What was actually earned is 477. 
Using resampling I found the average amount of dollars made per test was around 504.

# Confidence interval of Brownlows earning

<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows-->

```{r}
varSorted = sort(saveDollars)
LCL = varSorted[25]
UCL = varSorted[975]
print(paste("UCL = ",UCL,"LCL = ",LCL))
```


We assign values for mean, standard deviation, and sample size to rstudio in a formula.
We use our data to creat a code to provide us with a confidence interval.

```{r}
stdevDollars = sd(saveDollars)
meanDollars = mean(saveDollars)
CI = c(meanDollars-2*stdevDollars, meanDollars+2*stdevDollars)

print(CI)

```

Mean for losing is roughly 424.
Mean for winning is rougly 586. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->


```{r}
TotalDollars= 0
saveDollars=vector('numeric')
for(loop in 1:1000){
  
  TotalDollars= 0 
  ClassSize= 35
  for(i in 1:1137) {
    Bsamples = sample(Bdays,classSize, replace= TRUE)
    if (length(unique(Bsamples)) == classSize) {
      TotalDollars = TotalDollars -4.4
    } else{
      TotalDollars = TotalDollars +1
    }
  }
  saveDollars[loop] = TotalDollars +1
  }
  hist(saveDollars)
  var(saveDollars)
  print(paste("average return =",mean(saveDollars)))
```

We used the same set up as before and kept the original class size number. I tried payouts of -3, -10, and I found evntually that a payoff of -4.4 was the 'fair' amount. For every 4.4 dollars that he bets to every 1 dollar he breaks even.

```{r}
TotalDollars= 0
saveDollars=vector('numeric')
for(loop in 1:1000){
  
  TotalDollars= 0 
  ClassSize= 29
  for(i in 1:1137) {
    Bsamples = sample(Bdays,classSize, replace= TRUE)
    if (length(unique(Bsamples)) == classSize) {
      TotalDollars = TotalDollars -2
    } else{
      TotalDollars = TotalDollars +1
    }
  }
  saveDollars[loop] = TotalDollars +1
  }
  hist(saveDollars)
  var(saveDollars)
  print(paste("average return =",mean(saveDollars)))
  
```

We used the same set up as before and changed the class size numbers. By playing around with the numbers we changed the class size from 35 to 29. I tried class size of 15, 23, 25 and eventually found that the new class size is 29. For every -2 dollars that he bets to every 1 dollar he breaks even at a class size of 29.



